---
title: "Coffee Preferences"
subtitle: "Brilliant Cassowary - Appendix to report"
author:
  - "Nidhi Soma (ns848)"
  - "Joice Chen (jc3528)"
  - "Jinpeng Li (jl3496)"
  - "Stephen Syl-Akinwale (sis33)"
execute:
  echo: true
---

# Data cleaning

Please have tidyverse, tidymodels, usethis, and probably installed for our packages and libraries to work.

```{r}
#| label: loading packages
library(tidyverse)
library(tidyverse)
library(janitor)
library(dplyr)
library(skimr)
library(stringr)
library(ggplot2)
library(usethis)
library(tidymodels)
library(probably)
```

```{r}
#| label: loading-cleaned-data
coffee_df<-read_csv("data/GACTT_RESULTS_ANONYMIZED_v2.csv")

#remove NA columns
coffee_clean <- coffee_df |>
  select(-contains("flavorings")) |>
  select(-contains("Gender (please specify"))

#remove repetitive questions
coffee_clean <- coffee_clean |>
  mutate(`Where do you typically drink coffee?` = NULL) |>
  mutate(`How do you brew coffee at home?` = NULL)|>
  mutate(`On the go, where do you typically purchase coffee?` = NULL) |>
  mutate(`Do you usually add anything to your coffee?` = NULL) |>
  mutate(`What kind of diary do you add?` = NULL) |>
  mutate(`What kind of sugar or sweetener do you add?` = NULL) |>
  mutate(`Why do you drink coffee?` = NULL)

original_names <- colnames(coffee_clean)
tidy_names <- gsub(" ", "_", original_names)
tidy_names <- tolower(tidy_names)
tidy_names <- gsub("[[:punct:]]&&[^_]", "", tidy_names)

colnames(coffee_clean) <- tidy_names

#renaming
coffee_clean <- coffee_clean |>
  rename(
    age = "what_is_your_age?",
    cups_of_coffee_per_day = "how_many_cups_of_coffee_do_you_typically_drink_per_day?",
    how_else_at_home = "how_else_do_you_brew_coffee_at_home?",
    where_else_purchase_coffee = "where_else_do_you_purchase_coffee?",
    favorite_coffee_drink = "what_is_your_favorite_coffee_drink?",
    favorite_coffee = "please_specify_what_your_favorite_coffee_drink_is", 
    prefer_between_abc = "between_coffee_a,_coffee_b,_and_coffee_c_which_did_you_prefer?", 
    other_flavoring = "what_other_flavoring_do_you_use?", 
    best_described_before = "before_today's_tasting,_which_of_the_following_best_described_what_kind_of_coffee_you_like?", 
    like_coffee = "how_strong_do_you_like_your_coffee?", 
    roast_level = "what_roast_level_of_coffee_do_you_prefer?", 
    caffeine = "how_much_caffeine_do_you_like_in_your_coffee?", 
    own_coffee_expertise = "lastly,_how_would_you_rate_your_own_coffee_expertise?", 
    prefer_between_ad = "between_coffee_a_and_coffee_d,_which_did_you_prefer?", 
    favorite_overall_coffee = "lastly,_what_was_your_favorite_overall_coffee?", 
    time_spent_on_equipment = "approximately_how_much_have_you_spent_on_coffee_equipment_in_the_past_5_years?", 
    good_value_equipment = "do_you_feel_like_you’re_getting_good_value_for_your_money_with_regards_to_your_coffee_equipment?"
  )

colnames(coffee_clean) <- sapply(colnames(coffee_clean), function(name) {
  if (grepl("where_do_you_typically_drink_coffee", name)) {
    name <- gsub("where_do_you_typically_drink_coffee\\?_\\((.*)\\)", "drink_\\1", name)
  } else if (grepl("how_do_you_brew_coffee_at_home", name)) {
    name <- gsub("how_do_you_brew_coffee_at_home\\?_\\((.*)\\)", "at_home_\\1", name)
  } else if (grepl("on_the_go,_where_do_you_typically_purchase_coffee", name)) {
    name <- gsub("on_the_go,_where_do_you_typically_purchase_coffee\\?_\\((.*)\\)", "purchase_\\1", name)
  } else if (grepl("do_you_usually_add_anything_to_your_coffee", name)) {
    name <- gsub("do_you_usually_add_anything_to_your_coffee\\?_\\((.*)\\)", "add_to_\\1", name)
  } else if (grepl("what_kind_of_dairy_do_you_add", name)) {
    name <- gsub("what_kind_of_dairy_do_you_add\\?_\\((.*)\\)", "dairy_add_\\1", name)
  } else if (grepl("what_kind_of_sugar_or_sweetener_do_you_add", name)) {
    name <- gsub("what_kind_of_sugar_or_sweetener_do_you_add\\?_\\((.*)\\)", "sugar_sweetener_add_\\1", name)
  } else if (grepl("why_do_you_drink_coffee", name)) {
    name <- gsub("why_do_you_drink_coffee\\?_\\((.*)\\)", "reason_\\1", name)
  }
  name
}
)

#manually changing some more confusing names
coffee_clean_2 <- coffee_clean |>
  rename(at_home_coffee_brewing_machine = `at_home_coffee_brewing_machine_(e.g._mr._coffee)`,
         at_home_pod_or_capsule_machine = `at_home_pod/capsule_machine_(e.g._keurig/nespresso)`,
         at_home_coffee_extract = `at_home_coffee_extract_(e.g._cometeer)`,
         purchase_national_chain = `purchase_national_chain_(e.g._starbucks,_dunkin)`,
         add_to_none = `add_to_no_-_just_black`,
         add_to_milk = `add_to_milk,_dairy_alternative,_or_coffee_creamer`,
         sugar_sweetener_add_artificial_sweeteners = `sugar_sweetener_add_artificial_sweeteners_(e.g.,_splenda)`,
         sugar_sweetener_add_raw_sugar= `sugar_sweetener_add_raw_sugar_(turbinado)`,
         where_work = `do_you_work_from_home_or_in_person?`,
         monthly_coffee_cost = `in_total,_much_money_do_you_typically_spend_on_coffee_in_a_month?`,
         like_taste = `do_you_like_the_taste_of_coffee?`,
         know_where_coffee_from = `do_you_know_where_your_coffee_comes_from?`,
         most_pay = `what_is_the_most_you've_ever_paid_for_a_cup_of_coffee?`,
         most_willing_pay = `what_is_the_most_you'd_ever_be_willing_to_pay_for_a_cup_of_coffee?`,
         good_value_money = `do_you_feel_like_you’re_getting_good_value_for_your_money_when_you_buy_coffee_at_a_cafe?`) |>
  mutate(`what_kind_of_dairy_do_you_add?` = NULL)

#change type to categorical

coffee_clean_factors <- coffee_clean_2 |>
  mutate(age = factor(age),
         monthly_coffee_cost = factor(monthly_coffee_cost))|>
  mutate(across(like_taste:political_affiliation, factor)) |>
  mutate(across(like_coffee:caffeine, factor)) |>
  mutate(cups_of_coffee_per_day = as_factor(cups_of_coffee_per_day))|>
  mutate(best_described_before = factor(best_described_before))

#add category
coffee_clean_factors <- coffee_clean_factors |>
  mutate(age = fct_relevel(age, c("<18 years old",
                                  "18-24 years old",
                                  "25-34 years old",
                                  "35-44 years old", 
                                  "45-54 years old",
                                  "55-64 years old",
                                  ">65 years old")))|>
  mutate(monthly_coffee_cost = fct_relevel(monthly_coffee_cost, c(
    "<$20",
    "$20-$40",
    "$40-$60",
    "$60-$80",
    "$80-$100",
    ">$100"))) |>
  mutate(most_pay = fct_relevel(
    most_pay,
    c("Less than $2",
    "$2-$4",
    "$4-$6",
    "$6-$8",
    "$8-$10",
    "$10-$15",
    "$15-$20",
    "More than $20"
   ))) |>
  mutate(most_willing_pay = fct_relevel(
    most_willing_pay,
    c("Less than $2",
    "$2-$4",
    "$4-$6",
    "$6-$8",
    "$8-$10",
    "$10-$15",
    "$15-$20",
    "More than $20"
  ))) |>
  mutate(cups_of_coffee_per_day = fct_relevel(cups_of_coffee_per_day,
                                              c("Less than 1",
                                              "1",
                                              "2",
                                              "3",
                                              "4",
                                              "More than 4"))) |>
  mutate(caffeine = fct_relevel(caffeine,
                                c("Decaf", "Half caff", "Full caffeine"))) |>
  mutate(like_coffee = fct_relevel(like_coffee,
                                   c("Weak",
                                     "Somewhat light",
                                     "Medium",
                                     "Somewhat strong",
                                     "Very strong")))

coffee_clean_factors |>
  write_rds(file = "data/coffee_clean_factor.rds")

# coffee_remove_unused <- coffee_clean_factors |>
#   select(age, cups_of_coffee_per_day, add_to_none, add_to_milk, contains("sugar_sweetener_add"))

```

# Extra Data Analysis figures

```{r}
#| label: visual-all-prefs
#| fig-width: 8
#| fig-height: 6
#| 
#age vs  roast
#coffee like or not by age
roast_totals <- coffee_clean_factors |>
  group_by(roast_level) |>
  summarise(total = n())

age_roast_counts <- coffee_clean_factors |>
  filter(!is.na(roast_level) & !is.na(age)) |>
  group_by(age, roast_level) |>
  summarise(count = n()) |>
  left_join(roast_totals, by = "roast_level") |>
  mutate(percentage = count / total * 100)

ggplot(age_roast_counts, aes(x = roast_level, y = percentage, fill = age)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Coffee Roast Level Preferences by Age Group",
       x = "Roast Level",
       y = "Percentage") +
  theme_minimal() +
  scale_fill_viridis_d(end = 0.8)

# #Preferences by education level
coffee_clean_factors |>
  select(caffeine, education_level) |>
  drop_na() |>
  ggplot(aes(x = caffeine, fill = education_level)) +
  geom_bar() +
  labs(title = "Amount of caffeine in coffee by education level")

# Preference by Age
roast_totals <- coffee_clean_factors |>
  group_by(roast_level) |>
  summarise(total = n())
```

```{r}
#| label: extra sweetener graphs

#clean original dataset
sweet_edu_df <- coffee_clean_factors |>
  select(education_level, contains("sugar_sweetener_add")) |>
  drop_na()
names(sweet_edu_df) <- gsub("sugar_sweetener_add_", "", names(sweet_edu_df))
sweet_edu_df <- sweet_edu_df |> 
  mutate(education_level = case_when(
    education_level %in% c("Bachelor's degree", "Master's degree", "Doctorate or professional degree") ~ "bachelor",
    TRUE ~ "no_bachelor"
  )) |>
  mutate(education_level = factor(education_level,
                                  c("bachelor",
                                    "no_bachelor")))

#data set for analysis, with number of sugar types
sweet_edu_df_analysis <- sweet_edu_df |>
  rowwise() |>
  mutate(sweet_count = sum(across(granulated_sugar:raw_sugar), na.rm = TRUE))

#other dataset for visualization purposes
sweet_edu_df2 <- sweet_edu_df |>
  pivot_longer(
    cols = 2:9,
    names_to = "sweetener",
    values_to = "add"
  ) 

sweet_edu_df3 <- sweet_edu_df2 |>
  select(education_level, sweetener, add)|>
  filter(add == TRUE)

sweet_edu_df <- sweet_edu_df

#Popularity of sweeteners
sweet_edu_df2 |>
  ggplot(aes(y = sweetener, fill = add) ) +
  geom_bar()

#number of added sweeteners by education level
sweet_edu_df2 |>
  ggplot(aes(y = education_level), fill = "blue") +
  geom_bar() +
  labs(
    title = "Number of added sweeteners by education level",
    x = "Count",
    y = "Education level"
  )
```
These graphs show the distributions for each variable, so we can get an idea of the different groups we are working with. We see that granulated_sugar is the most popular sweetener, and people with higher than a Bachelor's degree are more common in this sample, which could potentially skew results. Next, we see how these two variables are related to each other through proportions, where we can more directly compare the visual differences.

```{r}
# #sweetener preference proportions by education level
# sweet_edu_df2 |>
#   ggplot(aes(y = sweetener, fill = add) ) +
#   geom_bar(position = "fill") +
#   facet_wrap(.~education_level)

#proportion preferring sweetener by education level
sweet_edu_df2 |>
  ggplot(aes(y = education_level, fill = add) ) +
  geom_bar(position = "fill") +
  facet_wrap(.~sweetener) +
  labs(
    title = "Proportion of people who add sweetener",
    subtitle= "By type of sweetener and education level",
    x = "Proportion",
    y= "Education level"
  )

```
