---
title: "Project title"
subtitle: "Exploratory data analysis"
execute:
  echo: true
---

```{r}
library(janitor)
library(dplyr)
library(skimr)
```

# Inserting the dataset on Coffee
```{r}
coffee_df<-read_csv("data/GACTT_RESULTS_ANONYMIZED_v2.csv")
```

# Research question(s)

Research question(s). State your research question (s) clearly.

# Data collection and cleaning

Have an initial draft of your data cleaning appendix. Document every step that takes your raw data file(s) and turns it into the analysis-ready data set that you would submit with your final project. Include text narrative describing your data collection (downloading, scraping, surveys, etc) and any additional data curation/cleaning (merging data frames, filtering, transformations of variables, etc). Include code for data curation/cleaning, but not collection.



# Data description

Have an initial draft of your data description section. Your data description should be about your analysis-ready data.

# Data limitations

Identify any potential problems with your dataset.

# Exploratory data analysis

Perform an (initial) exploratory data analysis.

# Questions for reviewers

List specific questions for your peer reviewers and project mentor to answer in giving you feedback on this phase.
