---
title: Coffee Preferences
author: Brilliant Cassowary <br> Jinpeng Li, Joice Chen, Nidhi Soma, Stephen Syl-Akinwale 
date: today
format:
  revealjs: 
    theme: simple
    scrollable: true
execute:
  echo: false
---

## Introduce the topic and motivation

Coffee is one of the most popular choices for students across campus. We wanted to see if there were any trends we could analyze about coffee preferences at large.

General research question: How does coffee preference and consumption habits differ by demographic attributes such as age, gender, education level, race, and political identity?

## Introduce the data

- Data collected from James Hoffman's "The Great American Coffee Taste Test"

- Surveyed almost 4000 respondents about their coffee preferences
- Participants were required weighing scale,a way of heating hot water, 
and cups that can accommodate about 250 mLs or eight ounces
- Coffee A - Light roast, Washed, Coffee B - Medium Roast, Coffee C - Dark roast, Coffee D - Fermented, Natural, Fruity

## Most Preferred Roast Level {.smaller}

Looking at these variables we can see age ranges 25-34 are the most avid coffee drinkers. We want to see more about their taste preferences.
```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(skimr)
library(stringr)
library(ggplot2)
library(usethis)
library(tidymodels)
library(probably)
coffee_clean_factors<- read_rds("data/coffee_clean_factor.rds")
#age vs  roast
#coffee like or not by age
roast_totals <- coffee_clean_factors |>
  group_by(roast_level) |>
  summarise(total = n())

age_roast_counts <- coffee_clean_factors |>
  filter(!is.na(roast_level) & !is.na(age)) |>
  group_by(age, roast_level) |>
  summarise(count = n()) |>
  left_join(roast_totals, by = "roast_level") |>
  mutate(percentage = count / total * 100)

ggplot(age_roast_counts, aes(x = roast_level, y = percentage, fill = age)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Coffee Roast Level Preferences by Age Group",
       x = "Roast Level",
       y = "Percentage") +
  theme_minimal() +
  scale_fill_viridis_d(end = 0.8)
```

## Similar Sweetener Preference by Educational Level {.smaller}
```{r}
#Education vs number of sweetners

sweet_edu_df <- coffee_clean_factors |>
  select(education_level, contains("sugar_sweetener_add")) |>
  drop_na()
names(sweet_edu_df) <- gsub("sugar_sweetener_add_", "", names(sweet_edu_df))
sweet_edu_df <- sweet_edu_df |> 
  mutate(education_level = case_when(
    education_level %in% c("Bachelor's degree", "Master's degree", "Doctorate or professional degree") ~ "bachelor",
    TRUE ~ "no_bachelor"
  )) |>
  mutate(education_level = factor(education_level,
                                  c("bachelor",
                                    "no_bachelor")))

sweet_edu_df2 <- sweet_edu_df |>
  pivot_longer(
    cols = 2:9,
    names_to = "sweetener",
    values_to = "add"
  ) 
sweet_edu_df2 |>
  ggplot(aes(y = sweetener, fill = add) ) +
  geom_bar(position = "fill") +
  facet_wrap(.~education_level) +
  labs(
    title = "Educational Level by Type of Sweetener",
    x = "Proportion",
    y= "Type of Sweetener",
    fill = "Adds Sweetener"
  )

```

## Hypothesis 1: Inference for Roast Level {.smaller}

```{r}
#| label: lin-reg-all-prefs
library(knitr)
  young_roast_reg<-  coffee_clean_factors |>
  select(age, roast_level) |>
  drop_na() |>
    mutate(roast_level_num = ifelse(roast_level == "Light" | roast_level == "Medium" | roast_level == "Nordic", 1, 0))
  
young_roast_fit <- linear_reg() |>
fit(roast_level_num ~ age, data = young_roast_reg)
kable(tidy(young_roast_fit))
```

- Model predicts that people aged 18-24 years old will have a 0.0229 lower proportion of people that prefer lighter roasts than people aged 25-34 years old, on average. 

- Leads to question: Are people aged 18-24 actually not the most likely group to prefer lighter roasts?

## Hypothesis 1: Visualizing Lighter/Darker Roasts {.smaller}

```{r}
#| label: show-independence-relationship
young_roast_binary <-young_roast_reg |>
  mutate(roast_level_binary = fct_collapse(
.f = roast_level,
Lighter = c("Light", "Medium", "Nordic"),
other_level = "Darker"))
  ggplot(young_roast_binary, mapping = aes(x = age, fill = roast_level_binary)) +
  geom_bar(position = "fill") +
    scale_fill_viridis_d() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

- If there were no relationship between these two variables, we would expect to see the purple bars reaching to the same height, regardless of age group. 

- Are differences just random noise or a meaningful relationship? - Chi square test!

## Hypothesis 1: Chi squared test for Roast Level

Null hypothesis: There is no association between age group and a lighter roast level preference.

Alternative hypothesis: There is an association between age group and a lighter roast level preference.

```{r}
#| label: chi-sqr-test
#source: https://infer.netlify.app/articles/chi_squared
# calculate the observed statistic
observed_indep_statistic <- young_roast_binary %>%
  specify(roast_level_binary ~ age) %>%
  hypothesize(null = "independence") %>%
  calculate(stat = "Chisq")

# generate the null distribution using randomization
null_dist_roast <- young_roast_binary %>%
  specify(roast_level_binary ~ age) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>%
  calculate(stat = "Chisq")
```
---
```{r}
# visualize the null distribution and test statistic!
null_dist_roast %>%
  visualize() + 
  shade_p_value(observed_indep_statistic,
                direction = "greater")
```

```{r}

# calculate the p value from the observed statistic and null distribution
p_value_independence <- null_dist_roast %>%
  get_p_value(obs_stat = observed_indep_statistic,
              direction = "greater")

kable(p_value_independence)

```
We can reject the null hypothesis because our p-value of approximately 0 is less than the alpha value of 0.05.

## Conclusions + future work {.smaller}

- For roast level, there is sufficient evidence that there is an association between age group and preferring lighter roasts of coffee. 
- Certain age groups are more prone to enjoying a certain taste, while others prefer the caffeine. 
- For number of sweeteners and educational level, we cannot really come to a conclusion

Future directions:

- Text analysis for the columns specifying people's reason for enjoying coffee versus their age group to explore this hypothesis.

- Make results more specific and relevant to Cornell community  - held back by limitations on next slide

## Limitations
- Answers were not mandatory, lots of absent values

- Only conducted in the US

- Respondents could have misrepresented their measurements

- Provided by Cometeer coffee company: possibly different tastes and limited selection

- Population is not representative: respondents are already avid coffee drinkers
