---
title: "Project title"
subtitle: "Brilliant Cassowary - Report"
author:
  - "Nidhi Soma (ns848)"
  - "Joice Chen (jc3528)"
  - "Jinpeng Li (jl3496)"
  - "Stephen Syl-Akinwale (sis33)"
echo: false
warning: false
message: false
---

```{r}
library(tidyverse)
library(tidymodels)
library(probably)
library(skimr)
```

# Introduction
One of the most widely known beverages in today's age is coffee. It is present in many different settings, ranging from students and employees to casual coffee shop frequenters and critical coffee enthusiasts.
Our current research questions are:
- How does self-perceived coffee experience affect actual and stated preferences?
- How does coffee preference and consumption habits differ by demographic
attributes such as age, gender, education level, race, and political identity?

# Data description
*What are the observations (rows) and the attributes (columns)?*

The observations represent an individual respondent to the survey. There are 4,042 rows. The columns are questions that they answered, ranging from demographic data to coffee preferences. There are 98 of these columns.

*Why was this dataset created?*

To understand the general public's preferences as consumers for coffee. Additionally, since Cometeer funded the creation of this dataset, they may be interested to know people's preferences in order to make their coffee capsules more appealing to a wider market.

*Who funded the creation of the dataset?*

World champion barista James Hoffmann and Cometeer – a subscription service that makes flash-frozen coffee capsules.

*What processes might have influenced what data was observed and recorded and what was not?*

The survey quickly was sold out, and Hoffman's audience in general is coffee specialists. That will likely skew the population surveyed to be people who likely prefer specialty coffee, so it may be a biased sample. Additionally, this survey was conducted through people ordering tasting kits online, which were then sent to the participants to prepare and complete voluntarily, so there may have been differences in that. One example is that because participants were following a livestream to demonstrate how to do their taste test, their coffees may have been out long enough to have cooled, which could be another unaccounted variable that affected their taste preferences.

*What preprocessing was done, and how did the data come to be in the form that you are using?*

Zip codes and geographic data seemed to have been removed. Participants were anonymized to protect their privacy. It wasn't disclosed how Hoffmann and his team collected all the taste test results that participants filled out, but once they got that data, they made it into a spreadsheet to be shared with the public.

*If people are involved, were they aware of the data collection and if so, what purpose did they expect the data to be used for?*

People were involved, and they were made aware of the data collection by a YouTube video stating that the purpose of this taste test was to understand coffee preferences in the USA. The participants had to order the coffee tasting kit on their own in order to participate, showing their willingness to accept these terms. Hoffmann also made his intentions clear in his video with why he wanted to collect the data, and that he was planning to publicize the raw data later on.

*Notes*

Coffee A - Light roast, Washed
Coffee B - Medium Roast
Coffee C - Dark roast
Coffee D - Fermented, Natural, Fruity

*Loading data*

```{r}
#| label: loading packages
library(tidyverse)
library(janitor)
library(dplyr)
library(skimr)
library(stringr)
library(ggplot2)
library(usethis)
library(tidymodels)
```
```{r}
#| label: loading-cleaned-data
coffee_df<-read_csv("data/GACTT_RESULTS_ANONYMIZED_v2.csv")

#remove NA columns
coffee_clean <- coffee_df |>
  select(-contains("flavorings")) |>
  select(-contains("Gender (please specify"))

#remove repetitive questions
coffee_clean <- coffee_clean |>
  mutate(`Where do you typically drink coffee?` = NULL) |>
  mutate(`How do you brew coffee at home?` = NULL)|>
  mutate(`On the go, where do you typically purchase coffee?` = NULL) |>
  mutate(`Do you usually add anything to your coffee?` = NULL) |>
  mutate(`What kind of diary do you add?` = NULL) |>
  mutate(`What kind of sugar or sweetener do you add?` = NULL) |>
  mutate(`Why do you drink coffee?` = NULL)

original_names <- colnames(coffee_clean)
tidy_names <- gsub(" ", "_", original_names)
tidy_names <- tolower(tidy_names)
tidy_names <- gsub("[[:punct:]]&&[^_]", "", tidy_names)

colnames(coffee_clean) <- tidy_names

#renaming
coffee_clean <- coffee_clean |>
  rename(
    age = "what_is_your_age?",
    cups_of_coffee_per_day = "how_many_cups_of_coffee_do_you_typically_drink_per_day?",
    how_else_at_home = "how_else_do_you_brew_coffee_at_home?",
    where_else_purchase_coffee = "where_else_do_you_purchase_coffee?",
    favorite_coffee_drink = "what_is_your_favorite_coffee_drink?",
    favorite_coffee = "please_specify_what_your_favorite_coffee_drink_is", 
    prefer_between_abc = "between_coffee_a,_coffee_b,_and_coffee_c_which_did_you_prefer?", 
    other_flavoring = "what_other_flavoring_do_you_use?", 
    best_described_before = "before_today's_tasting,_which_of_the_following_best_described_what_kind_of_coffee_you_like?", 
    like_coffee = "how_strong_do_you_like_your_coffee?", 
    roast_level = "what_roast_level_of_coffee_do_you_prefer?", 
    caffeine = "how_much_caffeine_do_you_like_in_your_coffee?", 
    own_coffee_expertise = "lastly,_how_would_you_rate_your_own_coffee_expertise?", 
    prefer_between_ad = "between_coffee_a_and_coffee_d,_which_did_you_prefer?", 
    favorite_overall_coffee = "lastly,_what_was_your_favorite_overall_coffee?", 
    time_spent_on_equipment = "approximately_how_much_have_you_spent_on_coffee_equipment_in_the_past_5_years?", 
    good_value_equipment = "do_you_feel_like_you’re_getting_good_value_for_your_money_with_regards_to_your_coffee_equipment?"
  )

colnames(coffee_clean) <- sapply(colnames(coffee_clean), function(name) {
  if (grepl("where_do_you_typically_drink_coffee", name)) {
    name <- gsub("where_do_you_typically_drink_coffee\\?_\\((.*)\\)", "drink_\\1", name)
  } else if (grepl("how_do_you_brew_coffee_at_home", name)) {
    name <- gsub("how_do_you_brew_coffee_at_home\\?_\\((.*)\\)", "at_home_\\1", name)
  } else if (grepl("on_the_go,_where_do_you_typically_purchase_coffee", name)) {
    name <- gsub("on_the_go,_where_do_you_typically_purchase_coffee\\?_\\((.*)\\)", "purchase_\\1", name)
  } else if (grepl("do_you_usually_add_anything_to_your_coffee", name)) {
    name <- gsub("do_you_usually_add_anything_to_your_coffee\\?_\\((.*)\\)", "add_to_\\1", name)
  } else if (grepl("what_kind_of_dairy_do_you_add", name)) {
    name <- gsub("what_kind_of_dairy_do_you_add\\?_\\((.*)\\)", "dairy_add_\\1", name)
  } else if (grepl("what_kind_of_sugar_or_sweetener_do_you_add", name)) {
    name <- gsub("what_kind_of_sugar_or_sweetener_do_you_add\\?_\\((.*)\\)", "sugar_sweetener_add_\\1", name)
  } else if (grepl("why_do_you_drink_coffee", name)) {
    name <- gsub("why_do_you_drink_coffee\\?_\\((.*)\\)", "reason_\\1", name)
  }
  name
}
)

#manually changing some more confusing names
coffee_clean_2 <- coffee_clean |>
  rename(at_home_coffee_brewing_machine = `at_home_coffee_brewing_machine_(e.g._mr._coffee)`,
         at_home_pod_or_capsule_machine = `at_home_pod/capsule_machine_(e.g._keurig/nespresso)`,
         at_home_coffee_extract = `at_home_coffee_extract_(e.g._cometeer)`,
         purchase_national_chain = `purchase_national_chain_(e.g._starbucks,_dunkin)`,
         add_to_none = `add_to_no_-_just_black`,
         add_to_milk = `add_to_milk,_dairy_alternative,_or_coffee_creamer`,
         sugar_sweetener_add_artificial_sweeteners = `sugar_sweetener_add_artificial_sweeteners_(e.g.,_splenda)`,
         sugar_sweetener_add_raw_sugar= `sugar_sweetener_add_raw_sugar_(turbinado)`,
         where_work = `do_you_work_from_home_or_in_person?`,
         monthly_coffee_cost = `in_total,_much_money_do_you_typically_spend_on_coffee_in_a_month?`,
         like_taste = `do_you_like_the_taste_of_coffee?`,
         know_where_coffee_from = `do_you_know_where_your_coffee_comes_from?`,
         most_pay = `what_is_the_most_you've_ever_paid_for_a_cup_of_coffee?`,
         most_willing_pay = `what_is_the_most_you'd_ever_be_willing_to_pay_for_a_cup_of_coffee?`,
         good_value_money = `do_you_feel_like_you’re_getting_good_value_for_your_money_when_you_buy_coffee_at_a_cafe?`) |>
  mutate(`what_kind_of_dairy_do_you_add?` = NULL)

#change type to categorical

coffee_clean_factors <- coffee_clean_2 |>
  mutate(age = factor(age),
         monthly_coffee_cost = factor(monthly_coffee_cost))|>
  mutate(across(like_taste:political_affiliation, factor)) |>
  mutate(across(like_coffee:caffeine, factor)) |>
  mutate(cups_of_coffee_per_day = as_factor(cups_of_coffee_per_day))|>
  mutate(best_described_before = factor(best_described_before))

#add category
coffee_clean_factors <- coffee_clean_factors |>
  mutate(age = fct_relevel(age, c("<18 years old",
                                  "18-24 years old",
                                  "25-34 years old",
                                  "35-44 years old", 
                                  "45-54 years old",
                                  "55-64 years old",
                                  ">65 years old")))|>
  mutate(monthly_coffee_cost = fct_relevel(monthly_coffee_cost, c(
    "<$20",
    "$20-$40",
    "$40-$60",
    "$60-$80",
    "$80-$100",
    ">$100"))) |>
  mutate(most_pay = fct_relevel(
    most_pay,
    c("Less than $2",
    "$2-$4",
    "$4-$6",
    "$6-$8",
    "$8-$10",
    "$10-$15",
    "$15-$20",
    "More than $20"
   ))) |>
  mutate(most_willing_pay = fct_relevel(
    most_willing_pay,
    c("Less than $2",
    "$2-$4",
    "$4-$6",
    "$6-$8",
    "$8-$10",
    "$10-$15",
    "$15-$20",
    "More than $20"
  ))) |>
  mutate(cups_of_coffee_per_day = fct_relevel(cups_of_coffee_per_day,
                                              c("Less than 1",
                                              "1",
                                              "2",
                                              "3",
                                              "4",
                                              "More than 4"))) |>
  mutate(caffeine = fct_relevel(caffeine,
                                c("Decaf", "Half caff", "Full caffeine"))) |>
  mutate(like_coffee = fct_relevel(like_coffee,
                                   c("Weak",
                                     "Somewhat light",
                                     "Medium",
                                     "Somewhat strong",
                                     "Very strong")))

```

# Data analysis

We address our first question on actual self-perceived coffee experience compared to stated preferences. We use examples of people stating their preferences and look at the difference in data where they have a taste test.


*Pre-registered hypotheses*

Our first pre-registered hypothesis was that younger people 18 - 25 prefer their coffee with lighter roasts.
```{r}
#| label: visual-hyp-prefs
# test <- coffee_clean_factors |>
#   select(age, favorite_coffee_drink, favorite_coffee, best_described_before, like_coffee, roast_level, prefer_between_abc, prefer_between_ad, favorite_overall_coffee)

  coffee_clean_factors |>
  select(age, roast_level) |>
  drop_na() |>
  filter(age == "18-24 years old") |>
  ggplot(aes(x = roast_level, fill = age)) +
  geom_bar() +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Roast Level Preferences",
    subtitle = "For respondents aged 18-24 only",
    x= "Roast level",
    y = "Frequency"
  )
```
We see that there is a clear peak around the Light and Medium roast levels, suggesting that in this sample, younger people that are 18-24 years old have a strong preference for Light and Medium roast levels. However, there is more of a preference for Dark roasts than Nordic roasts, even though Nordic roasts are also generally considered to be lighter roasts.

This leads to the question of whether the younger age range of 18-24 is the most likely age to prefer lighter roast levels (Light, Medium, or Nordic). We investigate this below by first visualizing the preferences of all the age ranges, and then using a linear regression model. 
```{r}
#| label: visual-all-prefs
#| fig-width: 8
#| fig-height: 6
coffee_clean_roast <- coffee_clean %>%
  mutate(roast_level = ifelse(is.na(roast_level), "No Preference", roast_level))

ggplot(coffee_clean_roast, aes(x = age, fill= roast_level)) +
  geom_bar(position = "dodge") +
  labs(title = "Preferred Coffee Roast Levels",
       x = "Age Group",
       y = "Frequency",
       fill= "Roast preference") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
 scale_fill_viridis_d()
```
In the future, exploring a logistic regression model may also lead to more accurate results, since the data when visualized doesn't appear to be very linear.
```{r}
#| label: lin-reg-all-prefs
  young_roast_reg<-  coffee_clean_factors |>
  select(age, roast_level) |>
  drop_na() |>
    mutate(roast_level_num = ifelse(roast_level == "Light" | roast_level == "Medium" | roast_level == "Nordic", 1, 0))
  
young_roast_fit <- linear_reg() |>
fit(roast_level_num ~ age, data = young_roast_reg)
tidy(young_roast_fit)

# glance(young_roast_fit)$adj.r.squared
#for diff of props, the output is saying one group is % more probable than the other
#topic modeling for text-analysis?
```
Each of the coefficients gives the proportion of people choosing lighter roasts for the corresponding age group relative to the reference level of <18 years old. The reference level has a coefficient of 0, meaning its estimated proportion is the intercept, 0.842. Notably, the model predicts that people aged 18-24 years old will have a 0.029 lower proportion of people that prefer lighter roasts than people aged 25-34 years old, on average. So, the model predicts that younger people aged 18-24 are actually not the most likely group to prefer lighter roasts.


*Pre-registered Hypothesis 2*
Our second pre-registered hypothesis is was that people with a bachelor's degree will higher would be more likely to add more types of sweeteners to their coffee as compared to those without a bachelor's degree

```{r}
#| label: 2data-visualization
#clean original dataset
sweet_edu_df <- coffee_clean_factors |>
  select(education_level, contains("sugar_sweetener_add")) |>
  drop_na()
names(sweet_edu_df) <- gsub("sugar_sweetener_add_", "", names(sweet_edu_df))
sweet_edu_df <- sweet_edu_df |> 
  mutate(education_level = case_when(
    education_level %in% c("Bachelor's degree", "Master's degree", "Doctorate or professional degree") ~ "bachelor",
    TRUE ~ "no_bachelor"
  )) |>
  mutate(education_level = factor(education_level,
                                  c("bachelor",
                                    "no_bachelor")))

#data set for analysis, with number of sugar types
sweet_edu_df_analysis <- sweet_edu_df |>
  rowwise() |>
  mutate(sweet_count = sum(across(granulated_sugar:raw_sugar), na.rm = TRUE))

#other dataset for visualization purposes
sweet_edu_df2 <- sweet_edu_df |>
  pivot_longer(
    cols = 2:9,
    names_to = "sweetener",
    values_to = "add"
  ) 

sweet_edu_df3 <- sweet_edu_df2 |>
  select(education_level, sweetener, add)|>
  filter(add == TRUE)

sweet_edu_df <- sweet_edu_df
  
#Popularity of sweeteners
sweet_edu_df2 |>
  ggplot(aes(y = sweetener, fill = add) ) +
  geom_bar()

#number of added sweeteners by education level
sweet_edu_df2 |>
  ggplot(aes(y = education_level), fill = add, color = "blue") +
  geom_bar()

#sweetener preference proportions by education level
sweet_edu_df2 |>
  ggplot(aes(y = sweetener, fill = add) ) +
  geom_bar(position = "fill") +
  facet_wrap(.~education_level)

#proportion preferring sweetener by education level
sweet_edu_df2 |>
  ggplot(aes(y = education_level, fill = add) ) +
  geom_bar(position = "fill") +
  facet_wrap(.~sweetener)

#number of sugar type by education
sweet_edu_df_analysis |>
  ggplot(aes(x = sweet_count, fill = education_level)) +
  geom_bar(position = "dodge")




```
The visualizations seem to show little difference between the distribution of the number of sugar types preferred by those with a Bachelor's or higher degree compared to that preferred by those with less than  Bachelor's degree. We can conduct a hypothesis test to further test our hypothesis. We conduct a linear regression below

```{r}
#| label: log-reg
edu_sweetener_fit1 <- logistic_reg() |>
  fit(education_level ~ sweet_count, data = sweet_edu_df_analysis)
tidy(edu_sweetener_fit1)


# edu_sweetener_fit2 <- logistic_reg() |>
#   fit(education_level ~ granulated_sugar, data = sweet_edu_df)
# tidy(edu_sweetener_fit)

```
The logistical regression model calculated the coefficients representing the that a person with less than a Bachelor's degree would on average have 0.06 less probability to choose less sugar additive types than a person with a Bachelor's degree or higher would, holding all else constant.

# Evaluation of significance

*Pre-registered Hypothesis 1*

We can investigate if the difference in proportions of preferring a lighter roast between age groups is statistically significant or not with a chi-squared test of independence.


```{r}
#| label: show-independence-relationship
young_roast_binary <-young_roast_reg |>
  mutate(roast_level_binary = fct_collapse(
.f = roast_level,
Lighter = c("Light", "Medium", "Nordic"),
other_level = "Darker"))
  ggplot(young_roast_binary, mapping = aes(x = age, fill = roast_level_binary)) +
  geom_bar(position = "fill") +
    scale_fill_viridis_d()
```

If there were no relationship between these two variables, we would expect to see the purple bars reaching to the same height, regardless of age group. We need to conduct a chi-squared test of independence to see if the differences we see here are just random noise or a meaningful relationship.

Null hypothesis: There is no association between age group and a lighter roast level preference.

Alternative hypothesis: There is an association between age group and a lighter roast level preference.

```{r}
#| label: chi-sqr-test
#source: https://infer.netlify.app/articles/chi_squared
# calculate the observed statistic
observed_indep_statistic <- young_roast_binary %>%
  specify(roast_level_binary ~ age) %>%
  hypothesize(null = "independence") %>%
  calculate(stat = "Chisq")

# generate the null distribution using randomization
null_dist_roast <- young_roast_binary %>%
  specify(roast_level_binary ~ age) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>%
  calculate(stat = "Chisq")

# visualize the null distribution and test statistic!
null_dist_roast %>%
  visualize() + 
  shade_p_value(observed_indep_statistic,
                direction = "greater")

# calculate the p value from the observed statistic and null distribution
p_value_independence <- null_dist_roast %>%
  get_p_value(obs_stat = observed_indep_statistic,
              direction = "greater")

p_value_independence

```

*Hypothesis 2*

We can investigate if the difference in accepted sugar additive types between education groups is statistically significant through a hypothesis test.

Null hypothesis: The true average number of types of sweetener preferred is the same between people in America with a Bachelor's degree or higher and people without a Bachelor's degree.



$$ H_0: \mu_{\text{bach}} - \mu_{\text{no bach}} = 0 $$

$$ H_a: \mu_{\text{bach}} - \mu_{\text{no bach}} \neq 0 $$

To evaluate whether the number of types of sweetener accepted differ between people with different education degrees, we will conduct a two-side hypothesis test

Alternative hypothesis: The true average number of types of sweetener preferred is not the same between people in America with a Bachelor's degree or higher and people without a Bachelor's degree.


```{r}
#| label: 2sample-statistics
sweet_edu_df_analysis |>
  group_by(education_level) |>
   summarize(mean = mean(sweet_count))
```
The observed statistic is 1.455-1.408 = 0.047

```{r}
#Hypothesis testing

ob_sweet <- sweet_edu_df_analysis |>
  specify(response = sweet_count, explanatory = education_level) |>
  hypothesize(null = "independence") |>
  calculate(stat = "diff in means")

# null distribution
null_dist_sweet <- sweet_edu_df_analysis |>
  specify(response = sweet_count, explanatory = education_level) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means")

# visualize the null distribution and test statistic
null_dist_sweet |>
  visualize() + 
  shade_p_value(ob_sweet,
                direction = "greater")

# calculate the p value from the observed statistic and null distribution
p_value <- null_dist_sweet |>
  get_p_value(obs_stat = ob_sweet,
              direction = "greater")

p_value
```


*Research Question 2*



# Interpretation and conclusions

*Hypothesis 1*
From the evaluation of significance, using an alpha value of 0.05, because our p-value of approximately 0 is less than 0.05, we can reject the null hypothesis. There is sufficient evidence that there is an association between age group and preferring lighter roasts of coffee.

In the data analysis, a possible explanation for younger people showing more of a preference for Dark roasts than Nordic roasts despite showing an overwhelming preference for Light and Medium roasts is that more people may be unfamiliar with what a Nordic roast is. For people with general coffee knowledge, seeing light, medium, and dark roasts tends to be more common than Nordic; however, we should still consider that this sample likely has many respondents that are knowledgeable about coffee. So, these results still may have another underlying cause and reveal something about younger people's coffee preferences.


# Limitations


# Acknowledgments

